### 2011-03-20 extra/sliver_pldi11/introduction.tex

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
supporting abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction increases.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis,
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
The general theme of all these approaches is that the
complexity of the abstraction is increased only in parts deemed relevant using
some feedback from a client query.

% Pruning idea: general
In this paper, we use {\em pruning}, a different approach from
the above selected refinement techniques for making static analyses scale.
Our approach works on any analysis that can be expressed as a Datalog program.
A Datalog program takes a set of input tuples (determined by the abstraction)
and derives new tuples via a set of inference rules; the answer to a client
query corresponds to whether a designated tuple can be derived.  The key idea
is to identify input tuples which are provably irrelevant for deriving the
query tuple and prune these away.  Selected refinement techniques
attempt to keep the set of input tuples small by not refining some of them; we
keep the set small by removing some of them entirely.

% Soundness
Pruning can be a dangerous affair though.  With selected refinement, at any
point we are still performing static analysis with respect to an abstraction
and therefore have soundness guarantees.  However, once we start pruning input
tuples, we are no longer running a valid static analysis and thus do not
automatically inherit soundness guarantees.  Nonetheless, we prove that our
method is {\em sound with respect to the client}.

% Completeness
While soundness is trivial for selected refinement but requires some argument for pruning,
the situation is reversed for {\em completeness}, that is, the guarantee that
an analysis is as precise as some target.  For selected refinement,
it is difficult to argue that the chosen coarser abstraction is as precise as if we
had refined everything.  However, with pruning, we are working directly with
the more complex abstraction, and thus automatically inherit its completeness
guarantees.  By removing input tuples, we can only prove more queries.

% Refinement algorithm
%To capitalize on the idea of {\em pruning}, we propose an algorithm, which we
%call Refinement-Pruning (\PR).  The idea is to start with a coarse abstraction
%and prune the input tuples before moving to a more expensive abstraction.

% Slivers
We apply our pruning technique to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects in the heap are abstracted using chains of
allocation sites.  Pruning in this setting corresponds to maintaining a subset
of these chains.  Although the main contribution of the paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise because we are pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction
than before due to pruning.  Second, we introduce an abstraction which limits
chains to length $k$ and also truncates to avoid having too many repeating
sites; this is an effective way to increase $k$ further (possible now due to
pruning) without getting bogged down by recursion.  Third, we show that an
abstraction that replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having a precise pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
algorithm is able to go well beyond $k=10$.

>>>>

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
underlying abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction is refined.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis;
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
We refer to these techniques as {\em selected refinement},
where the main idea is to only refine an abstraction along components
deemed relevant according to client feedback.

% Pruning setup
In this paper, we introduce {\em pruning}, a new approach which represents a
significant departure from existing selected refinement techniques.  Pruning is
applicable to static analyses expressed as a set of inference rules where a
program property of interest (a client query) is proven by the inability to
derive a designated fact using the given rules.  For concreteness, assume that
the static analysis is expressed as a Datalog program.  A Datalog program takes
a set of input tuples and derives new tuples via a set of inference rules.
These inference rules capture the abstract semantics of the static analysis and
the evaluation of the client query using the analysis result; the input tuples encode the program we are
analyzing and the abstraction we are using.  The program property is proven if the
designated query tuple cannot be derived.

% Pruning idea
The key idea behind pruning is to identify input tuples which are provably
irrelevant for deriving the query tuple and remove these tuples completely from
analysis.  Consequently, when the abstraction is refined, only the relevant
tuples are refined, resulting in potentially major computational savings.  It
is helpful to think of pruning in terms of program slicing, where
irrelevant parts of the program (irrelevant input tuples) are removed,
resulting in a smaller program that is cheaper to analyze.  Existing selected
refinement techniques attempt to keep the set of input tuples small by simply
not refining some of them; pruning keeps the set small by removing some of them
entirely.

% Soundness
Pruning can be a dangerous affair though.  With selected refinement,
we are always performing a static analysis with respect to an abstraction
and therefore inherit the soundness guarantees of abstract interpretation.
However, once we start pruning input tuples, we are no longer running a valid
static analysis of the original program.  Soundness therefore is no longer automatic,
though we do prove that our method is sound with respect to the given client.

% Completeness
While soundness is trivial for selected refinement but requires some
argument for pruning, the situation is reversed for {\em completeness}.  By
completeness, we mean that the analysis is as precise as
if we had refined all the components of an abstraction.  Selected
refinement only refines a subset of an abstraction, so it is unclear
that the resulting abstraction is as precise as an abstraction obtained by
refining all components.  However, with pruning, we conceptually work with the
fully-refined abstraction; and by removing input tuples, we cannot prove fewer
queries; thus, completeness is automatic.

% Refinement algorithm
To capitalize on the idea of pruning, we propose an algorithm, which we
call the {\em Prune-Refine algorithm}.  The idea is to start with a coarse abstraction
and prune the irrelevant input tuples before refining the abstraction; the
algorithm iterates until the query is proven or a specified computational budget
is reached.  We prove that the Prune-Refine algorithm computes the same answers to client queries
as directly using a refined abstraction without pruning, which would be precise
but possibly infeasible.

% Slivers
We apply pruning to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects in the heap are abstracted using chains of
allocation sites; these chains are the input tuples we maintain.
Although the main contribution of this paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise due to pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction.
Second, we introduce a new abstraction which limits
chains to length $k$ and truncates to avoid repeating
allocation sites; this is an effective way to increase $k$ without getting bogged down by long chains created by recursion.
Third, we show that an abstraction that
replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having a precise pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our Prune-Refine algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In a few cases, the non-pruning approaches hit a wall around $k=3$ but the Prune-Refine
algorithm is able to go well beyond $k=10$.
